A Relaxation Approach to Feature Selection for Mixed-Effects Models

### 2022-07-04 presentation/general_presentation.tex
Finish first pass for presentation

\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}
\newcommand{\ouralgo}{\ensuremath{\mathcal{MSR}3}}
\title{Feature Selection for Mixed-Effects Models}
\date{\today}
\author{Aleksei Sholokhov}
% \titlegraphic{\hfill\includegraphics[height=1.5cm]{logo.pdf}}

@@

\begin{frame}{Plan}
	\tableofcontents
\end{frame}

\section{Feature Selection for Linear Mixed-Effect Models}
\subsection{Linear Mixed-Effects Models}
\begin{frame}{Linear Mixed-Effect Models}

@@

TODO: add picture

@@

\begin{frame}{SR3-Relaxation for Mixed-Effect Models}
Original problem $\mathcal{FS-LME}$:
\eq{
	\min_{\beta \in \R^p,\, \gamma \in \R^{q}_+} \LL(\beta, \gamma) + R(\beta, \gamma)
}
Relaxed problem $\ouralgo$:
\eq{
	\label{eq:msr3_formulation_explicit}
	\min_{\beta, \tbeta \in \R^p,\, \gamma, \tgamma \in \R^{q}_+} \LL(\beta, \gamma) + \phi_\mu(\gamma) + \kappa_\eta(\beta - \tbeta, \gamma - \tgamma) + R(\beta, \gamma)
}
where the \textit{relaxation} $\kappa_\eta$ decouples the likelihood and the regularizer 
\eq{
	\kappa_\eta(\beta - \tbeta, \gamma - \tgamma) := \frac{\eta}{2}\|\beta - \tbeta\|^2_2 + \frac{\eta + \bar\eta}{2}\|\gamma - \tgamma\|_2^2
}
and the \textit{projection function} $\phi_\mu$ replaces $\gamma \geq 0$ with a log-barrier 
\eq{
	\phi_\mu(\gamma) := \begin{cases}
		-\mu\sum_{i=1}^q \ln(\gamma_i/\mu), & \mu > 0 \\
		\delta_{\R_+^q}(\gamma), & \mu = 0 \\
		+\infty, & \mu < 0
	\end{cases}
} 
\end{frame}

@@

\textbf{Key observation}: in practice, we don't need accurate solutions for~(\ref{eq:value_function_definition}): a few Newton iterations is enough.
\end{frame}

\begin{frame}{Value Function Reformulation}
TODO: insert image here
\end{frame}

@@

\textbf{Lemma:} For every $(\mu,\eta) \in \R_+\times\R_{++}$,
\eq{
	(\hat\beta,\hat\gamma) = \argmin_{(\beta,\gamma)}\LL_{\eta,\mu}((\beta, \gamma),(\tbeta, \tgamma)) \\
	\text{(equivalent to)} \\
	\exists \hat{v} \in \R_{+}^q \text{ s.t. } G_{\nu,\eta}((\beta,\gamma,\hat{v}),(\tbeta,\tgamma)) = 0
}
If $\mu > 0$, then $\hat{v} = -\nabla\phi_\mu(\hat{\gamma})$, and if $\mu = 0$, then $\hat{v}$ is the unique KKT multiplier associated with the constraint $0 \leq \gamma$.
\end{frame}

\begin{frame}{$\ouralgo$ Algorithm}
\input{../extras/08Pseudocode}
\end{frame}

@@

\begin{frame}{Performance in Comparison to Other Algorithms}
\begin{itemize}
	\item \textbf{Scenario 1}: $n=30$, $n_i=5$, $p = 9$, $q = 4$, with true parameters $\beta = (1, 1, 0, \dots, 0)$ and the covariance matrix $\Gamma$ being: 
		\eq{
			\Gamma = \begin{bmatrix}
				9 & 4.8 & 0.6 & 0 \\
				4.8 & 4 & 1 & 0 \\
				0.6 & 1 & 1 & 0\\
				0 & 0 & 0 & 0
			\end{bmatrix}
		}
	\item \textbf{Scenario 2}: everything as in Scenario 1, but $n=60$ and $n_i=10$.
\end{itemize}
\textbf{Competitors}:
\begin{itemize}
	\item \textbf{ALASSO:} 2 stage: A-LASSO+Newton and A-LASSO+PCO
	\item \textbf{M-ALASSO:} Adaptive LASSO + EM Algorithm
	\item \textbf{SCAD-P:} SCAD + Proxy Matrix for $\Gamma$
	\item \textbf{rPQL:} Quasi-Likelihood + Adaptive LASSO (for GLMMs)
\end{itemize}
\end{frame}

\begin{frame}{Performance in Comparison to Other Algorithms}
\begin{table}[H]
\begin{center}
\begin{tabular}{|l|l|c|c|c|c|c|}
\hline
Setup & Algoritm & \% C & \% CF & \% CR & MSE & TIME \\
\hline 
\hline
$n = 30$, $n_i = 5$ & \textbf{\ouralgo} & 58 & 72 & 78 & 0.66 & 0.015 \\
& rPQL & 88 & 98 & 88 & 0.88 & 26-59 \\ 
& M-ALASSO & 71 & 73 & 79 & - & -\\
& ALASSO & 79 & 81 & 96 & - & - \\
& SCAD-P & - & 90 & 86 & - & - \\
\hline 
$n = 60$, $n_i = 10$ & \textbf{\ouralgo} & 98 & 100  & 98 & 0.69& 0.018 \\
& rPQL & 98 & 99 & 98 & 0.97 &  26-59 \\ 
& M-ALASSO & 83 & 83 & 89 & - & - \\
& ALASSO & 95 & 96 & 99 & - & -\\
& SCAD-P & 100 & 100 & 100 & - & - \\
\hline

\end{tabular}
\end{center}
\caption{\label{table:krishna_setup_results} Comparison of feature selection algorithms. \% CF -- percent of models where true fixed effects were identified correctly, \% CR -- percent of models where true random effects were identified correctly, \% C -- both fixed and random effects were identified correctly.}	
\end{table}

\begin{frame}{Scalability Experiment}
\begin{itemize}
	\item $n = 60$, $n_i = 10$
	\item $p = q \in [4, 7, 10, \dots, 90]$, 200 experiments for each.
	\item $X_i = Z_i$, columns are drawn form $\NN(0, \Psi)$ where
	\[
		\Psi = \begin{bmatrix}
			9 & 4.8 & 0.6 \\
			4.8 & 4 & 1 \\
			0.6 & 1 & 1 \\
		\end{bmatrix}
	\]
	\item 50\% random coordinate in $\beta$ are active
	\item 70\% of those are also active in $\gamma$
\end{itemize}
\end{frame}

\begin{frame}{Scalability Experiment}
		\begin{center}
			\includegraphics<+>[width=0.8\textwidth]{Figures/scalability_accuracy}
			%\includegraphics<+>[width=0.8\textwidth]{Figures/scalability_mse}
		\end{center}
\end{frame}

\subsection{Application to Real-World Problems}

\begin{frame}{Contact Rate Modeling for COVID-19 Forecasting}
	\begin{itemize}
		\item $n = 60$ groups (countries and US states), $n_i \approx 50$
		\item $Y_i$ -- contact rate for COVID SEIIR Model 
		\item $p=q=5$ covariates related to temperature, mobility, population, testing; plus intercept.
	\end{itemize}
\end{frame}

\begin{frame}{Contact Rate Modeling for COVID-19 Forecasting}
	\begin{center}
		\includegraphics<+>[width=\textwidth]{Figures/fit_Alaska}
		\includegraphics<+>[width=\textwidth]{Figures/fit_Slovenia}
		\includegraphics<+>[width=\textwidth]{Figures/fit_Switzerland}
		\includegraphics<+>[width=\textwidth]{Figures/fit_Turkey}
	\end{center}
\end{frame}

\begin{frame}{Burden of Anxiety and Depression as Result of Bullying}
	\begin{itemize}
		\item $m = 10$ cohort studies, $n = 77$, highly unbalanced 
		\item $p = 13$, $q = 12$ (\texttt{time} was preselected fixed-only)
		\item Covariates are related to studies' designs.
	\end{itemize}
\end{frame}

\begin{frame}{Burden of Anxiety and Depression as Result of Bullying}
	\begin{center}
		\includegraphics[width=\textwidth]{Figures/bullying_data.csv_inclusion}
		%\includegraphics<+>[width=\textwidth]{Figures/bullying_data.csv_fixed_feature_selection}
		%\includegraphics<+>[width=\textwidth]{Figures/bullying_data.csv_random_feature_selection}

	\end{center}
\end{frame}
\section{Future Work}
\begin{frame}{Future Work: Theory}
\begin{theorem}[Conditions for Convergence to True Estimator]
	Under certain conditions the method converges in a finite number of iterations to $(\hat\beta, \hat\gamma)$ which projections $(\tbeta, \tgamma)$ belong to a $k$- and $j$-subspaces respectively that contain the true minimum $(\beta^*,\gamma^*)$.
\end{theorem}
\begin{theorem}[Consistency of Estimator]
	There exists a local minimizer $(\hat{\beta}, \hat{\gamma})$ for the proposed loss function, such that it is asymptotically consistent with true minimum $(\beta^*, \gamma^*)$. 
\end{theorem}
\begin{theorem}[Consistency in Zeros]
	If some coordinates of the true minimizer ($\beta^*, \gamma^*$) are zero, then it is also zero in $(\hat{\beta}, \hat{\gamma})$, given that the later is sufficiently close to the former.
\end{theorem}
\begin{theorem}[Asymptotic Normality]
	The proposed estimator $(\hat{\beta}, \hat{\gamma})$ asymptotically normally distributed around true minimizer ($\beta^*, \gamma^*$) in its true non-zero $k+j$-subspace.
\end{theorem}	
\end{frame}

\begin{frame}{Future work: Algorithm}
	\textbf{Question:} Will exponential smoothing of projection improve the accuracy? 
	{\small
	\begin{algorithm}[H]
	$\lambda_\beta = 0$; $\lambda_\gamma = 0$ \\
	\Repeat{$\tbeta \approx \beta, \tgamma \approx \gamma$}{
	$\lambda_\beta \leftarrow 2(1 + \lambda_\beta)$ \\
	$\lambda_\gamma \leftarrow 2(1 + \lambda_\gamma)$\\
	\Repeat{converges}{
		$\tbeta^{(k+1)} \leftarrow {\color{red}\delta}\proj_{\|\beta\|_0 \leq k}(\beta^{(k)}) + {\color{red}(1-\delta)\tbeta^{(k)}}$\\
		$\tgamma^{(k+1)} \leftarrow {\color{red}\delta}\proj_{\|\gamma\|_0 \leq s}(\gamma^{(k)}) + {\color{red}(1-\delta)\tgamma^{(k)}}$\\
		$\beta^{(k+1)}, \gamma^{(k+1)} \leftarrow \argmin_{\gamma \geq 0, \beta}\LL(\beta, \gamma) + \frac{\lambda_\beta}{2}\|\beta - \tbeta^{(k)}\|^2_2 + \frac{\lambda_\gamma}{2}\|\gamma - \tgamma^{(k)}\|_2^2$
	}
	}
	\BlankLine
	\end{algorithm}
	}
	\end{frame}

\begin{frame}{Future Work: Implementation}
		Can we increase $\lambda_\beta$ and $\lambda_\gamma$ in a more careful way to avoid potential stacking? 
		The approach can be based on the theorem:
		\begin{theorem}[Distance Between Minima]
	For a fixed dataset $(X_i, Y_i)$ and relaxation parameters $\lambda_\beta$, $\lambda_\gamma$ the distance between $(\beta^*, \gamma^*)$, the unconstrained minimizers of relaxed problem, and their projections $(\tbeta^*, \tgamma^*)$ is bounded by a constant $M$ depending on $(X_i, Y_i)$ and the relaxation parameters.
		\end{theorem}
\end{frame}

\begin{frame}{The End}
	\begin{center}
		\large Thank you for your attention!
	\end{center}
\end{frame}

\begin{frame}[allowframebreaks]{References}

  \bibliography{bibliography}
  \bibliographystyle{alpha}

>>>>

\usepackage{xspace}
\usepackage{caption}
\usepackage{subcaption}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}
\newcommand{\ouralgo}{\ensuremath{\mathcal{MSR}3}}
\title{A Relaxation Approach to Feature Selection for Mixed-Effects Models}
\date{\today}
\author{Aleksei Sholokhov, James V. Burke, Peng Zheng, and Aleksandr Aravkin}
% \titlegraphic{\hfill\includegraphics[height=1.5cm]{logo.pdf}}

@@

\section{Relaxed for Linear Mixed-Effect Models}
\subsection{Linear Mixed-Effects Models}
\begin{frame}{Linear Mixed-Effect Models}

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.25\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/l0_regularizer.pdf}
         \caption{$\ell_0$}
     \end{subfigure}%
     \begin{subfigure}[b]{0.25\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/l1_regularizer.pdf}
         \caption{$\ell_1$}
     \end{subfigure}%
     \begin{subfigure}[b]{0.25\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/lh_regularizer.pdf}
         \caption{$\ell_p,\, p=1/2$}
     \end{subfigure}%
     \begin{subfigure}[b]{0.25\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/scad_regularizer.pdf}
         \caption{SCAD}
     \end{subfigure}
     \caption{Three simple graphs}
     \label{fig:three graphs}
\end{figure}

@@

\begin{frame}{SR3-Relaxation for Mixed-Effect Models}
Original problem $\mathcal{FS-LME}$:
\eq{
	\min_{\beta \in \R^p,\, \gamma \in \R^{q}_+} \LL(\beta, \gamma) + R(\beta, \gamma)
}
Relaxed problem $\ouralgo$:
\eq{
	\label{eq:msr3_formulation_explicit}
	\min_{\beta, \tbeta \in \R^p,\, \gamma, \tgamma \in \R^{q}_+} \LL(\beta, \gamma) + \phi_\mu(\gamma) + \kappa_\eta(\beta - \tbeta, \gamma - \tgamma) + R(\tbeta, \tgamma)
}
where the \textit{relaxation} $\kappa_\eta$ decouples the likelihood and the regularizer 
\eq{
	\kappa_\eta(\beta - \tbeta, \gamma - \tgamma) := \frac{\eta}{2}\|\beta - \tbeta\|^2_2 + \frac{\eta + \bar\eta}{2}\|\gamma - \tgamma\|_2^2
}
and the \textit{projection function} $\phi_\mu$ replaces $\gamma \geq 0$ with a log-barrier 
\eq{
	\phi_\mu(\gamma) := \begin{cases}
		-\mu\sum_{i=1}^q \ln(\gamma_i/\mu), & \mu > 0 \\
		\delta_{\R_+^q}(\gamma), & \mu = 0 \\
		+\infty, & \mu < 0
	\end{cases}
} 
\end{frame}

@@

\textbf{Key observation}: in practice, we don't need accurate solutions for~(\ref{eq:value_function_definition}): a few Newton iterations keep the solution close to the central path.
\end{frame}

\begin{frame}{Value Function Reformulation}
\begin{figure}
	\includegraphics[width=0.7\textwidth]{Figures/intuition_current.pdf}
	\caption{\label{fig:intuition_sr3} Comparison of the level-sets for the original likelihood (left) and $\ouralgo$-likelihood (right), for fixed (top) and random (bottom) effects.}
\end{figure}
\end{frame}

@@

\textbf{Lemma:} For every $(\mu,\eta) \in \R_+\times\R_{++}$,
\eq{
	&(\hat\beta,\hat\gamma) = \argmin_{(\beta,\gamma)}\LL_{\eta,\mu}((\beta, \gamma),(\tbeta, \tgamma)) \\
	& \iff \\
	& \exists \hat{v} \in \R_{+}^q \text{ s.t. } G_{\nu,\eta}((\beta,\gamma,\hat{v}),(\tbeta,\tgamma)) = 0
}
If $\mu > 0$, then $\hat{v} = -\nabla\phi_\mu(\hat{\gamma})$, and if $\mu = 0$, then $\hat{v}$ is the unique KKT multiplier associated with the constraint $0 \leq \gamma$.
\end{frame}

\begin{frame}{$\ouralgo$-fast Algorithm}
\input{../extras/08Pseudocode}
\end{frame}

@@

\begin{frame}{Application to Synthetic Problems}
\textbf{The Experiment} 
\begin{itemize}
	\item The number of fixed effects $p$ and random effects $q$ is 20.
	\item $\beta = \gamma = \frac{1}{2}[1,2,3,\dots,10, 0\dots,0]$
	\item 9 groups with sizes [10, 15, 4, 8, 3, 5, 18, 9, 6]
	\item $X_i \sim \NN(0, I)^p$, $Z_i = X_i$, $\varepsilon_i \sim \NN(0, 0.3^2I)$
	\item Each experiment is repeated 100 times.
	\item Grid-search for $\eta \in [10^{-4}, 10^{2}]$, golden search for $\lambda \in [0, 10^5]$
	\item Final model is chosen to maximize BIC
\end{itemize}
\begin{table}
	\input{../extras/performance_table_short_current}
\end{table}
\end{frame}

\begin{frame}{Application to Synthetic Problems}
\begin{figure}
	\includegraphics[width=\textwidth]{Figures/performance_picture_current.pdf}
\end{figure}
Benefits:
\begin{itemize}
	\item $\ouralgo$-relaxation has similar (and sometimes better!) feature selection performance than the original likelihood.
	\item $\ouralgo$-fast optimization accelerates the compute time by $\sim 10^2$.
\end{itemize}
Setbacks:
   \begin{itemize}
   	\item $\eta$ is a new hyperparameter to tune.
   \end{itemize}
\end{frame}

\begin{frame}{$\ell_0$-based Covariate Selection for Bullying Study from GBD}
	\begin{figure}
		\includegraphics[width=\textwidth]{Figures/bullying_data_assessment_selection.pdf}
		\caption{Fixed and random covariate selection for Bullying dataset from~\cite{GBD}. The model selected 9 covariates, 7 of which were historically significant, and did not select 4 covariates, 1 of which was historically significant.}
	\end{figure}
\end{frame}

\begin{frame}{Thank You!}
	The code is available on GitHub: \\
	\href{github.com/aksholokhov/pysr3}{https://github.com/aksholokhov/pysr3}
	\begin{itemize}
		\item All estimators are fully compatible to \texttt{sklearn} library.
		\item Implements SR3 for linear, generalized-linear, and linear mixed-effect models.
		\item Has tutorials, tests, and documentation.
	\end{itemize}

	\textbf{References}: 
	\bibliographystyle{plain}
	\bibliography{bibliography.bib}
\end{frame}
